Glue-on AdS holography for $T\bar T$-deformed CFTs

### 2023-06-20 alpha/GlueonAdS/glueon-beamer.tex
Tune glueon slides for physics audience

\documentclass[aspectratio=169,10pt
%	,draft
%	,handout
	,noamsthm
]{beamer}

@@

\title{Glue-on AdS holography for $T\bar T$-deformed CFTs}
\subtitle{Holography duality beyond AdS/CFT}
\author{%
	Wen-Xin Lai 
	\texstringonly{\,}\textkai{赖文昕}%
	\texstringonly{\\[1ex]
		\textit{\small in collaboration with} \\
			Luis Apolo,
			Peng-Xiang Hao \textkai{郝鹏翔}
			and Wei Song \textkai{宋伟}\\[2ex]
		\sidenote{\arxiv{2303.04836}}
		\ submitted to \textit{JHEP}
	}%
}
\institute{\small%
	Yau Mathematical Sciences Center,
	Tsinghua\texstringonly{
%		\\\large\ccbyncsajp
	}%
}
\date{}

@@

\begin{document}
{%%% TITLEPAGE
\setbeamercolor{title in head/foot}{
		use=palette quaternary
		,fg=palette quaternary.bg
	}
\logo{}
\begin{frame}
%	\titlegraphic{\vspace{2\baselineskip}}
	\titlepage
	\tikz [remember picture,overlay]
	\node at ([
		yshift=-2.5\baselineskip,xshift=.3\linewidth
	] current page.center) {
		\begin{minipage}{.5\textwidth}
		\centering
		\includegraphics[height=.45\textheight]{img/escher_links.png}
		\\[2ex]
		\footnotesize fig:~\textcite{Dunham:Escher} on Escher's art
		\hspace{-1em}
		\end{minipage}
	};

\end{frame}
}%%% TITLEPAGE

@@

\section{\textbf{Background:} AdS/CFT holography}

\newcommand{\stateAdsCft}{\ensuremath{
	\textrm{String Theory on $\mrm{AdS}_{d+1}$ background}
	&\ \equiv\ 
	\textrm{Conformal Field Theory $\mrm{CFT}_{d}$}
\\
	\textrm{Asymptotically $\mrm{AdS}_{d+1}$ Gravity}
	&\ \equiv\ 
	\textrm{Large $N$ $\mrm{CFT}_{d}$ at asympt.~boundary}
}}

\begin{frame}{AdS/CFT --- holography duality}{%
	\textcite{Maldacena:1997re} --- \citetitle{Maldacena:1997re}%
}
\vspace{-1\baselineskip}
\begin{align}
\stateAdsCft
\end{align}
\vspace{-1.5\baselineskip}
\pause
\begin{itemize}
\item (asympt.) AdS: Anti-de Sitter spacetime, geometry with negative Ricci curvature $R < 0$, solution of Einsteins equations with cosmological constant $\Lambda < 0$:
\begin{equation}
	R_{\mu\nu} - \frac{1}{2}\,R g_{\mu\nu}
	+ \Lambda g_{\mu\nu}
	= 8\pi G_N T_{\mu\nu}^{\textit{matter}},\quad
	\Lambda < 0
\label{eq:einstein}
\end{equation}

\pause
\item CFT: conformal field theory, quantum field theory (QFT) invariant under rescaling\textsuperscript{*}, well defined \& well understood in math
\begin{equation}
	g_{ij} \sim e^{2\Omega(x)} g_{ij}
\end{equation}
*\,``ruler-independent'', {\footnotesize unless there is an \textit{``anomaly''}, but everything is still under control}
\end{itemize}
\vspace{\baselineskip}
\end{frame}

\begin{frame}{AdS/CFT --- a model of quantum gravity}{%
	\textcite{Maldacena:1997re} -- \citetitle{Maldacena:1997re}%
}
\vspace{-1.5\baselineskip}
\begin{align*}
\stateAdsCft
\end{align*}
\vspace{-1.5\baselineskip}
\pause
\begin{itemize}

\item A model of \textsc{quantum gravity}, discovered from dualities in \textsc{string theory}
	\begin{itemize}
	\item Dualities: equivalent descriptions for the same theory, physicist's ``proof'' of AdS/CFT
	\item No general proof yet in the mathematical sense, but there is \textit{overwhelming} evidence
	\end{itemize}

\pause
\item Universal / ubiquitous
\begin{itemize}
	\item Model agnostic: $\mrm{AdS}_5\times S^5$, $\mrm{AdS}_3\times S^3 \times T^4$, symmetric orbifolds, minimal models, ``monsters'', ... \textit{\small Maldacena, Witten, GKP, ABJM, Gaberdiel, and many many more}
	\item Our work: based on universal properties of $\mrm{AdS}_3/\mrm{CFT}_2$ $(d=2)$.

@@

\newcommand{\figAdsCft}{
\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/ads-cft.png}

	\vspace{-.3\baselineskip}
	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{figure}
}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- gravity trapped in a \textit{Poincar\'e cylinder}}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}
\begin{columns}
\begin{column}{.45\textwidth}
\figAdsCft
\end{column}
\begin{column}{.55\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric,
	\begin{equation}
	\hspace{-2em}
		ds^2 = \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j \label{fggauge}
	\end{equation}

@@

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Aharony:1999ti} -- \citetitle{Aharony:1999ti}
}
\begin{columns}
\begin{column}{.4\textwidth}
\figAdsCft
\end{column}
\begin{column}{.6\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric for $\rho > 0$, \\
	\textcite{Banados:1998gg}
	\begin{equation}
	\hspace{-2.5em}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ \label{banados}
	\end{equation}
	$g^{(2)} \leadsto \mathcal L(u)$, $\bar{\mathcal L}(v)$: arbitrary periodic functions,\\
	where $u,v = \varphi \mp it$, $du\,dv = d\varphi^2 + dt^2$
%	\vspace{-.5\baselineskip}

\pause
	\item A consequence of Einstein equations \eqref{eq:einstein} \\
	in $(2+1)$ dimensions.\\
	\textit{Too many constraints, too little freedom!}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Banados:1992wn} -- \citetitle{Banados:1992wn}
}
\begin{columns}
\begin{column}{.4\textwidth}
\figAdsCft
\end{column}
\begin{column}{.6\textwidth}
	\begin{itemize}
	\item Different \textbf{geometries} in the ``bulk''
	\begin{equation}
	\hspace{-2.5em}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ \label{banados}
	\end{equation}
	specified by $\mathcal L(u)$, $\bar{\mathcal L}(v)$: arbitrary periodic functions

	\vspace{.5\baselineskip}

\pause
	\item ... maps to different \textbf{states} in the boundary $\mrm{CFT}_2$
	\begin{itemize}
		\item \textit{state:} excitations of the boundary field theory
		\item \textit{spectrum} of the Hamiltonian operator, $\hat{H}$
	\end{itemize}
%	\item In a \textit{holographic} theory, $,\\
%	a boundary state can be prepared by a bulk path integral.
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- the holographic dictionary}{%
	\textcite{Aharony:1999ti} -- \citetitle{Aharony:1999ti}
}
\begin{columns}
\begin{column}{.4\textwidth}
\figAdsCft
\end{column}
\begin{column}{.6\textwidth}
	\begin{itemize}
	\item \textit{Dictionary:} bulk $\leftrightarrow$ boundary quantities

	\begin{itemize}
	\item spectrum,
	\item symmetry,
	\item correlations, ...
	\end{itemize}

	\item Partition function:
	\begin{equation}
		\mcal{Z}_{\pd B}
		= \sum_{i \in \textit{states}} \ave{e^{-\beta \hat{H}}}_i
		= \mcal{Z}_{B\textit{ulk}}
	\end{equation}
	weighted sum over the spectrum

@@

\begin{frame}{\TTbar deformations --- motivations}{%
	Comments taken from \textcite{Cui:2023jrb}
}
\begin{columns}
\begin{column}{.45\textwidth}
\figAdsCft
\end{column}
\begin{column}{.55\textwidth}
	\begin{itemize}
	\item Ideally: AdS/CFT
	\item Reality: non-AdS space and non-CFT
	\begin{itemize}
		\item possible to extend holographic principle to more general context?
		\item attempt: \textbf{\textit{deformations}} on both sides\\
		in a controllable way
	\end{itemize}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{\TTbar deformations --- definition}{%
	\textcite{Zamolodchikov:2004ce},\\
	revived by \textcite{Smirnov:2016lqw,Cavaglia:2016oda}
}
\begin{itemize}
\item Path integral formalism:
\begin{equation}
	\mcal{Z}
	= \sum_{i \in \textit{states}} \ave{e^{-\beta \hat{H}}}_i
	= \int \DD{\phi(x)} e^{-I[\phi(x)]}
\label{eq:partition}
\end{equation}
$I[\phi(x)]$: action (energy) functional of the fields $\phi(x)$\\
\phantom{$I[\phi(x)]$: }captures the dynamics of the theory

@@

\item Stress tensor $T_{ij}[\phi(x)]$: energy momentum, local and covariant version of $\hat{H}$
\begin{equation}
%	\label{TTbardefintro}
	\partial_\mu I = -{8 \pi} \int d^2x \, T\bar T =  -{\pi} \int d^2x\,\big( T^{ij}T_{ij}- (T^i_i)^2 \big)
\end{equation}
For $\mrm{CFT}_2$,\ $T\bar{T} = T_{xx} T_{\bar{x}\bar{x}}$,\ where $x, \bar{x} = \varphi' \mp it'$.
{\footnotesize We've set the length scale $\ell = 1$.}

@@

\begin{frame}{\TTbar deformations --- ``solvable''}{%
	\textcite{Zamolodchikov:2004ce},\\
	revived by \textcite{Smirnov:2016lqw,Cavaglia:2016oda}
}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Define of \TTbar as the flow of action:
\begin{equation}
	\label{TTbardefintro}
	\partial_\mu I = -{8 \pi} \int d^2x \, T\bar T =  -{\pi} \int d^2x\,\big( T^{ij}T_{ij}- (T^i_i)^2 \big)
\end{equation}
This is a \textit{flow equation} from some generic QFT specified by $I(\mu = 0)$.

\item \textbf{``Solvable'':} the deformed spectrum of $\hat{H}(\mu)$ and $\hat{J}(\mu)$ on a cylinder of radius $R$\\
can be solved, and it's a simple function of the undeformed $E(0)$, $J(0)$:
\begin{equation}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0) % \label{ttbarspectrum}
\end{equation}
under simple (and reasonable) assumptions (e.g.~translation invariance et al).

\begin{frame}{\TTbar deformations --- properties}{%
	\textcite{Zamolodchikov:2004ce}\\
	\citetitle{Zamolodchikov:2004ce}
}
\vspace{-1.\baselineskip}
\begin{equation}
	\partial_\mu I = -{8 \pi} \int d^2x\,T\bar T
	\label{TTbardef}
\end{equation}
\begin{equation}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0)  \label{ttbarspectrum}
\end{equation}
\vspace{-\baselineskip}
\begin{itemize}
\item It seems that the \TTbar deformation is well-defined non-perturbatively for generic $\mrm{QFT}_2$,\\
yet mysterious: computations of many quantities are difficult\\
\begin{itemize}
\item e.g.~correlators: \textcite{Kraus:2018xrn,Cardy:2019qao,Cui:2023jrb}
\item ``asymptotic fragility'': \textcite{Dubovsky:2017cnj}
\end{itemize}

@@

\newcommand{\figCutoffAds}{
\begin{figure}[!h]
	\hfill\includegraphics[width=.8\linewidth]{img/cutoff-ads.jpg}

	\centering
	\vspace{-.3\baselineskip}
	\caption{Cutoff $\mrm{AdS}_3/\TTbar$ duality}

	\vspace{-.3\baselineskip}
	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{figure}
}

\begin{frame}{Cutoff $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\begin{column}{.4\textwidth}
\figCutoffAds
\end{column}
\begin{column}{.6\textwidth}
	\begin{itemize}
	\item Question: what is the AdS dual of \TTbar deformation?\\

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\begin{column}{.35\textwidth}
\figCutoffAds
\end{column}
\begin{column}{.65\textwidth}
	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}

@@

	\item Passed many non-trivial tests
	\item Generalized to ``mixed boundary conditions''\\
	by \textcite{Guica:2019nzm}

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}
\begin{columns}
\begin{column}{.35\textwidth}
\figCutoffAds
\end{column}
\begin{column}{.65\textwidth}
\vspace{-.5\baselineskip}
	\begin{itemize}
	\item \textbf{Significance:} gravity inside a finite box\\
		with hard Dirichlet walls

@@

	Luis Apolo,
	Peng-Xiang Hao \textkai{郝鹏翔}
	and Wei Song \textkai{宋伟}\\[2ex]
\sidenote{\arxiv{2303.04836}}
\ submitted to \textit{JHEP}

>>>>

\documentclass[
	10pt
%	,draft
	,handout
	,noamsthm
]{beamer}

@@

\title{Glue-on AdS holography for $T\bar T$-deformed CFTs}
\subtitle{An extended AdS\,/\,\TTbar duality \textit{(beyond infinity)}}
\author{%
	Wen-Xin Lai 
	\texstringonly{\,}\textkai{赖文昕}%
	\texstringonly{
		\textit{\small in collaboration with} \\
			Luis Apolo,
			Peng-Xiang Hao \textkai{郝鹏翔}
			and Wei Song \textkai{宋伟}\\[2ex]
		\sidenote{\arxiv{2303.04836}}%
%		\ submitted to \textit{JHEP}
	}%
}
\institute{\small%
	Yau Mathematical Sciences Center,\\
	Tsinghua\texstringonly{
%		\\\large\ccbyncsajp
	}%
}
\date{}

@@

\begin{document}
{%%% TITLEPAGE
\setbeamercolor{title in head/foot}{
		use=palette quaternary
		,fg=palette quaternary.bg
	}
\logo{}
\begin{frame}
%	\titlegraphic{\vspace{2\baselineskip}}
	\titlepage
	\tikz [remember picture,overlay]
	\node at ([
		yshift=-4.8\baselineskip,
		xshift=.26\linewidth
	] current page.center) {
		\begin{minipage}{.5\textwidth}
		\centering
		\includegraphics[height=.33\textheight]{img/smbc-cantor-cropped.png}
		\\[.2ex]
		\footnotesize \url{https://smbc-comics.com/comic/cantor}
		\hspace{-1.5em}
		\end{minipage}
	};
\end{frame}
}%%% TITLEPAGE

@@

\section{\textbf{Warmup:} $\mrm{AdS}_{d+1}$/$\mrm{CFT}_{d}$ at $d = 2$}

\newcommand{\stateAdsCft}{
\begin{align*}
	\textrm{Strings on $\mrm{AdS}_{d+1}$ background}
	&\ \equiv\ 
	\textrm{Conformal Field Theory $\mrm{CFT}_{d}$}
\\
	\textrm{\textit{asympt.}~$\mrm{AdS}_{d+1}$ Gravity}
	&\ \equiv\ 
	\textrm{Large $N$ $\mrm{CFT}_{d}$ at \textit{asympt.}~boundary}
\end{align*}
}

\newcommand{\citeMaldacena}{%
	\textcite{Maldacena:1997re} --- \scriptsize\citetitle{Maldacena:1997re}
}

\begin{frame}{AdS/CFT --- holography duality}{%
	\citeMaldacena%
}
\vspace{-2.5\baselineskip}
\stateAdsCft
\vspace{-1.8\baselineskip}
\begin{figure}[!h]
	\centering
	\includegraphics[width=.5\linewidth]{img/ads-cft.png}

	\vspace*{-.8\baselineskip}
	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{figure}
\vspace{-1.3\baselineskip}
\end{frame}

%\begin{equation}
%	R_{\mu\nu} - \frac{1}{2}\,R g_{\mu\nu}
%	+ \Lambda g_{\mu\nu}
%	= 8\pi G_N T_{\mu\nu}^{\textit{matter}},\quad
%	\Lambda < 0
%\label{eq:einstein}
%\end{equation}

\begin{frame}{AdS/CFT --- a model of quantum gravity}{%
	\citeMaldacena
}
\vspace{-2.4\baselineskip}
\stateAdsCft
\vspace{-1.6\baselineskip}
\pause
\begin{itemize}

\item A model of \textsc{quantum gravity}
\item ... originating from \textsc{string theory}
\item ... seemingly \textsc{Universal / ubiquitous}

\pause
\begin{itemize}
	\item Model agnostic: $\mrm{AdS}_5\times S^5$, $\mrm{AdS}_3\times S^3 \times T^4$, symmetric orbifolds, minimal models, ``monsters'', ... \textit{\small Maldacena, Witten, GKP, ABJM, Gaberdiel, and many many more}
	\item Our work: pure Einstein gravity in $\mrm{AdS}_3$ / $\mrm{CFT}_2$ $(d=2)$.

@@

%%% LaTeX3e!
\NewDocumentCommand{\figAdsCft}{ O{.28} }{
\begin{column}{#1\textwidth}
\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/ads-cft-cleaner.png}

	\vspace{-.01\baselineskip}
	\scriptsize\textcite{AldegundePWSep22}\ \,
\end{figure}
\end{column}
}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- gravity trapped in a \textit{Poincar\'e cylinder}}{%
	\textcite{Aharony:1999ti}
}
\begin{columns}
\figAdsCft
\begin{column}{.75\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric, $\rho \sim z^2 \sim r^{-2}$
	\begin{equation}
	\hspace{-2em}
		ds^2 = \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j \label{fggauge}
	\end{equation}

@@

\newcommand{\eqBanados}{
	\begin{equation*}
	\hspace{-2em}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ \,
	\end{equation*}
}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Banados:1992wn}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric for $\rho > 0$, \\
	\textcite{Banados:1998gg}
	\eqBanados
	$g^{(2)} \leadsto \mathcal L(u)$, $\bar{\mathcal L}(v)$: arbitrary periodic functions,\\
	where $u,v = \varphi \mp it$, $du\,dv = d\varphi^2 + dt^2$
%	\vspace{-.5\baselineskip}

\pause
	\item A consequence of Einstein equations \\
	in $(2+1)$ dimensions.\\
	\textit{Too many constraints, too little freedom!}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Aharony:1999ti}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item Different \textbf{geometries} in the ``bulk''
	\eqBanados
	specified by $\mathcal L(u)$, $\bar{\mathcal L}(v)$: periodic functions, maps to different \textbf{states} in the boundary $\mrm{CFT}_2$ with stress energy $\ave{T(u)}_\psi$, $\ave{\bar{T}(v)}_\psi$

	\item \textit{Dictionary:} bulk $\leftrightarrow$ boundary quantities:\\
		\begin{itemize}
		\item spectrum,
			symmetry,\\
			partition functions,
			correlators, ...
		\end{itemize}
	\end{itemize}

\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}{\TTbar deformations --- motivations}{%
	Comments taken from \textcite{Cui:2023jrb}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item Ideally: AdS/CFT
	\item Reality: non-AdS space and non-CFT
	\begin{itemize}
		\item possible to extend holographic principle to more general context?
		\item attempt: \textbf{\textit{deformations}} on both sides\\
		in a controllable way
	\end{itemize}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\newcommand{\citeTTbar}{%
	\textcite{Zamolodchikov:2004ce}, revived by \textcite{Smirnov:2016lqw} and \textcite{Cavaglia:2016oda} \textit{et al}
}

\begin{frame}{\TTbar deformations --- ``solvable''}{%
	\citeTTbar
}
\begin{itemize}

\item Define ``\,\TTbar\,'' as the flow of action:
\begin{equation}
	\partial_{\sidenote{\mu}} I = -{8 \pi} \int d^2x \, T\bar T =  -{\pi} \int d^2x\,\big( T^{ij}T_{ij}- (T^i_i)^2 \big)
	\label{TTbardef}
\end{equation}
For $\mrm{CFT}_2$,\ $T\bar{T} = T_{xx} T_{\bar{x}\bar{x}}$,\ where $x, \bar{x} = \varphi' \mp it'$.

\item The stress tensor $T_{ij}(\sidenote{\mu})$ on the right hand side flows with the deformation parametrized by $\sidenote{\mu}$.

@@

\item This is thus a \textit{differential equation} for a flow, starting from some generic QFT specified by $I(\mu = 0)$. In this work we shall start from $\mrm{CFT}_2$, but note that \TTbar is \textit{irrelevant}: conformal symmetry will be broken.

@@

\begin{frame}{\TTbar deformations --- ``solvable''}{%
	\citeTTbar
}
\begin{itemize}

\item \textbf{``Solvable'':} the deformed spectrum of $\hat{H}(\mu)$ and $\hat{J}(\mu)$ on a cylinder of radius $R$ can be solved, 
and it's a simple function\\
 of the undeformed $E(0)$, $J(0)$:
\begin{equation}
\hspace{-2em}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0) % \label{ttbarspectrum}
\end{equation}
under simple (and reasonable) assumptions (e.g.~translation invariance et al).

\item \textbf{Variants:} suppose the theory has some ``components'' labeled by $w$,

\begin{itemize}
\item double-trace $\TTbar = (\sum_w T_w) (\sum_w \bar{T}_w)$ (this work)
\item single-trace $\TTbar' = \sum_w (T_w \bar{T}_w)$ (e.g.~\textit{\citeauthor{Cui:2023jrb}},~23)
\end{itemize}

\end{itemize}

\vspace{.5\baselineskip}
\end{frame}

\begin{frame}{\TTbar deformations --- properties}{%
	\textcite{Zamolodchikov:2004ce}\\
	\citetitle{Zamolodchikov:2004ce}
}
\vspace{-1.\baselineskip}
\begin{equation}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0)  \label{ttbarspectrum}
\end{equation}
\vspace{-\baselineskip}
\begin{itemize}
\item It seems that the \TTbar deformation is well-defined non-perturbatively for generic $\mrm{QFT}_2$,
yet it's still mysterious and difficult\\
\begin{itemize}
\item e.g.~correlators: \textcite{Kraus:2018xrn,Cardy:2019qao,Cui:2023jrb}
\item ``asymptotic fragility'': \textcite{Dubovsky:2017cnj}
\end{itemize}

@@

\newcommand{\figCutoffAds}[1][.3]{
\begin{column}{#1\textwidth}
\begin{figure}[!h]
	\hfill\includegraphics[width=\linewidth]{img/cutoff-ads.jpg}

	\vspace{-.5\baselineskip}
	\flushleft
	\ \scriptsize\textcite{AldegundePWSep22}
\end{figure}
\end{column}
\hspace{-2em}
}

\begin{frame}{Cutoff $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-2em}
\begin{column}{.75\textwidth}
	\begin{itemize}
	\item Question: what is the AdS dual of \TTbar deformation?\\

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.7\textwidth}
	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}

@@

	\item Passed many non-trivial tests
	\item Related to ``mixed boundary conditions''\\
	by \textcite{Guica:2019nzm}

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.75\textwidth}
\vspace{-.5\baselineskip}
	\begin{itemize}
	\item \textbf{Significance:} gravity inside a finite box\\
		with hard Dirichlet walls

@@

	Luis Apolo,
	Peng-Xiang Hao \textkai{郝鹏翔}
	and Wei Song \textkai{宋伟}\\[2ex]
\sidenote{\arxiv{2303.04836}}
%\ submitted to \textit{JHEP}
